Approximation as Differentiation: Notes

### 2025-05-25 main/introduction.tex
elliott18.

Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be rewound or replayed after the fact, to compute the Galois connection described above. It would be satisfying to have a denotational account of \GPS as well, for which traces would then offer a plausible implementation technique (perhaps one of several). As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08}, which points to a way to develop such an approach. We have already hinted at this in the description above, but let us now make it explicit.

>>>>

Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be rewound or replayed after the fact, to compute the Galois connection described above. It would be satisfying to have a denotational account of \GPS as well, for which traces would then offer a plausible implementation technique (perhaps one of several). As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18}, which points to a way to develop such an approach. We have already hinted at this in the description above, but let us now make it explicit.
